---
title: "FINAL_Data_Clean"
output: html_document
date: "2024-12-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




# 1. Load packages

```{r}
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(cowplot)
```

# 2. Load data

```{r}
injury_data <- read.csv('all_data/NBA_Injury.csv')
All_RegSs_Per_Games <- read.csv('all_data/NBA_All_Seasons_Regular_Season_Per_Games.csv')
All_RegSs_Totals <- read.csv('all_data/NBA_All_Seasons_Regular_Season_Totals.csv')
```

```{r}
player_data <- read.csv('NBA_player_data/player_data.csv')
players <- read.csv('NBA_player_data/players.csv')
Seasons_Stats <- read.csv('NBA_player_data/Seasons_Stats.csv')

```


# 3. Data Clean: Injury Data

```{r}
# Check the contents of the 'Notes' column
head(injury_data$Notes, 10)
```
## 3.1. Key Words

### 3.1.1. Pre-processing of note

```{r}
# Extraction of all words
notes <- injury_data$Notes[!is.na(injury_data$Notes)]
words <- unlist(str_extract_all(tolower(notes), "\\b\\w+\\b"))
```

```{r}
# Check for duplicate elements
duplicated_elements <- words[duplicated(words)]
# Output duplicate elements
#unique(duplicated_elements)
```


```{r}
# summrize frequency of keywords
word_counts <- sort(table(words), decreasing = TRUE)
head(word_counts, 20)

```

### 3.1.2. Define injury type


```{r}
# Define two types of keywords
injury_methods <- list(
  "Sprains" = c("\\bsprain\\b", "\\bsprained\\b"),
  "Strains" = c("\\bstrain\\b", "\\bstrained\\b"),
  "Soreness" = c("\\bsore\\b"),
  "Fractures" = c("\\bfracture\\b", "\\bfractured\\b", "\\bbroken\\b"),
  "Overuse Injuries" = c("\\boveruse\\b", "\\btendinitis\\b", "\\bbursitis\\b", "\\bstress fracture\\b"),
  "Inflammation" = c("\\binflammation\\b", "\\bswelling\\b", "\\bitis\\b"),
  "Ligament Tears" = c("\\btear\\b", "\\bacl\\b", "\\bmcl\\b", "\\blcl\\b", "\\bpcl\\b"),
  "Fatigue" = c("\\bfatigue\\b", "\\bovertraining\\b", "\\bburnout\\b")
)

injury_locations <- list(
  "Back Issues" = c("\\bback\\b"),
  "Knee Issues" = c("\\bknee\\b"),
  "Tendon Issues" = c("\\bachilles\\b", "\\btendon\\b", "\\bachilles tendon\\b"),
  "Ankle Issues" = c("\\bankle\\b"),
  "Shoulder Issues" = c("\\bshoulder\\b"),
  "Wrist Issues" = c("\\bwrist\\b"),
  "Elbow Issues" = c("\\belbow\\b"),
  "Hand Issues" = c("\\bhand\\b", "\\bfinger\\b"),
  "Foot Issues" = c("\\bfoot\\b", "\\btoe\\b",  "\\bheel\\b"),
  "Head Issues" = c("\\bhead\\b", "\\bconcussion\\b"),
  "Hip Issues" = c("\\bhip\\b", "\\bgroin\\b", "\\bpelvis\\b"),
  "Neck Issues" = c("\\bneck\\b", "\\bcervical\\b", "\\bwhiplash\\b"),
  "Leg Issues" = c("\\bleg\\b", "\\bcalf\\b", "\\bthigh\\b", "\\bhamstring\\b", "\\bquad\\b"),
  "Arm Issues" = c("\\barm\\b", "\\bforearm\\b", "\\bbicep\\b", "\\btricep\\b")
)

# define other injuries
injuries_keywords <- c("\\binjury\\b", "\\binjuries\\b", "\\bminor injury\\b", "\\bgeneral injury\\b", "\\bhurt\\b")

```


### 3.1.3. Create new columns: Injuries tab

```{r}
# extract notes as a df
notes_with_labels <- data.frame(Notes = notes)  

# Initialize two new columns
notes_with_labels$Injury_Method <- NA
notes_with_labels$Injury_Location <- NA
# Initialize new column `Other_Injury` to NA
notes_with_labels$Other_Injury <- NA

# Marking the injury method
for (method in names(injury_methods)) {
  notes_with_labels$Injury_Method[is.na(notes_with_labels$Injury_Method) & 
    sapply(notes_with_labels$Notes, function(note) {
      any(grepl(paste(injury_methods[[method]], collapse = "|"), note, ignore.case = TRUE))
    })
  ] <- method
}

# Marking the injury location
for (location in names(injury_locations)) {
  notes_with_labels$Injury_Location[is.na(notes_with_labels$Injury_Location) & 
    sapply(notes_with_labels$Notes, function(note) {
      any(grepl(paste(injury_locations[[location]], collapse = "|"), note, ignore.case = TRUE))
    })
  ] <- location
}

# Check the unmarked records
unclassified_notes <- is.na(notes_with_labels$Injury_Method) & is.na(notes_with_labels$Injury_Location)

# Check for the keyword “Injuries” in Uncategorized Notes.
notes_with_labels$Other_Injury[unclassified_notes & 
  sapply(notes_with_labels$Notes, function(note) {
    any(grepl(paste(injuries_keywords, collapse = "|"), note, ignore.case = TRUE))
  })
] <- "Other"

# Mark unmatched rows as NA
notes_with_labels$Other_Injury[is.na(notes_with_labels$Other_Injury)] <- NA

# View Results
#head(notes_with_labels, 10)

```

```{r}
injury_data$Injury_Method <- notes_with_labels$Injury_Method
injury_data$Injury_Location <- notes_with_labels$Injury_Location
injury_data$Other_Injury <- notes_with_labels$Other_Injury

```

## 3.2. Visualization

### 3.2.1. Table

```{r}
# Frequency of Injury_Method and Injury_Location after exclusion of NA values
method_table <- table(notes_with_labels$Injury_Method, useNA = "no")  # summarize injury method
location_table <- table(notes_with_labels$Injury_Location, useNA = "no")  # summarize injury location
```


### 3.2.2. Histogram

```{r}
# Table of distribution of methods of injury
method_df <- as.data.frame(method_table)
colnames(method_df) <- c("Method", "Count")

# Plotting Histogram of the distribution of injury methods
method_plot <- ggplot(method_df, aes(x = reorder(Method, Count), y = Count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Horizontal bar chart
  geom_text(aes(label = Count), hjust = -0.2) +  # Labeling values at the top of bars
  expand_limits(y = max(method_df$Count) * 1.2) +  # x-axis range increased by 20%
  labs(title = "Distribution of Injury Methods", x = "Injury Method", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10))

# Table of Distribution of Injured Location
location_df <- as.data.frame(location_table)
colnames(location_df) <- c("Location", "Count")

# Plotting Histogram of the distribution of injury location
location_plot <- ggplot(location_df, aes(x = reorder(Location, Count), y = Count)) +
  geom_bar(stat = "identity", fill = "darkorange") +
  coord_flip() +  ## Horizontal bar chart
  geom_text(aes(label = Count), hjust = -0.2) +  # Labeling values at the top of bars
  expand_limits(y = max(location_df$Count) * 1.2) +  # x-axis range increased by 20%
  labs(title = "Distribution of Injury Locations", x = "Injury Location", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10))

# # Merge the two diagrams
# combined_plot <- plot_grid(
#   method_plot, location_plot,
#   ncol = 1,  # One column arrangement (top and bottom layout)
#   align = "v",  # vertical alignment
#   labels = c("A", "B")  # Adding subgraph labels
# )

print(method_plot)
print(location_plot)

```


## 3.3. Remove NA rows


```{r}
# Delete all rows with NA
injury_data_cleaned <- injury_data[!(is.na(injury_data$Injury_Method) &
                                     is.na(injury_data$Injury_Location) &
                                     is.na(injury_data$Other_Injury)), ]

#head(injury_data_cleaned, 10)  

```

## 3.4. Check existing df

### 3.4.1. View Injury Data


```{r}
# Inspection results
cat("Remaining rows:", nrow(injury_data_cleaned), "\n")

# Counts the number of rows where only Injury_Method has a value
only_method <- sum(!is.na(injury_data_cleaned$Injury_Method) & 
                   is.na(injury_data_cleaned$Injury_Location) & 
                   is.na(injury_data_cleaned$Other_Injury))

# Counts the number of rows where only Injury_Location has a value
only_location <- sum(is.na(injury_data_cleaned$Injury_Method) & 
                     !is.na(injury_data_cleaned$Injury_Location) & 
                     is.na(injury_data_cleaned$Other_Injury))

# Counts the number of rows where only Other_Injury has a value
only_other <- sum(is.na(injury_data_cleaned$Injury_Method) & 
                  is.na(injury_data_cleaned$Injury_Location) & 
                  !is.na(injury_data_cleaned$Other_Injury))

cat("The number of rows where only Injury_Method has a value:", only_method, "\n")
cat("The number of rows where only Injury_Location has a value:", only_location, "\n")
cat("The number of rows where only Other_Injury has a value:", only_other, "\n")
```

### 3.4.2. View Other_Injury

```{r}
# Filter rows where only Other_Injury has a value
only_other_rows <- notes_with_labels[
  !is.na(notes_with_labels$Other_Injury), 
]

only_other_rows
```

### 3.4.3. View 'Acquired' column

#### 3.4.3.1. Whether full NA

```{r}
all(is.na(injury_data_cleaned$Acquired))
```

#### 3.4.3.2. If all the same

```{r}
# Get the value of the first row
first_value <- injury_data_cleaned$Acquired[1]

# Determine if all values are the same as the first row
all_same <- all(injury_data_cleaned$Acquired == first_value, na.rm = TRUE)


if (all_same) {
  print("All values in this column are the same as the first one")
} else {
  print("The values in this column are not exactly the same")
}

```

#### 3.4.3.3. Remove all spaces

```{r}
# Remove spaces with gsub()
injury_data_cleaned$Acquired <- gsub(" ", "", injury_data_cleaned$Acquired)
table(injury_data_cleaned$Acquired)
```



```{r}
acquired_rows <- injury_data_cleaned[injury_data_cleaned$Acquired %in% c("LazarHayward", "ShaunLivingston"), ]
acquired_rows
```

#### 3.4.3.4. Delete these two rows & empty columns

```{r}
injury_data_cleaned <- injury_data_cleaned[!injury_data_cleaned$Acquired %in% c("ShaunLivingston", "LazarHayward"), ]
injury_data_cleaned$Acquired <- NULL
```

## 3.5. Reset line number

```{r}
rownames(injury_data_cleaned) <- NULL
injury_data_cleaned$X <- NULL
```


## 3.6. Processing date

```{r}
# 拆分 Date 列为 Year, Month, Day 三列
injury_data_cleaned <- injury_data_cleaned %>%
  separate(Date, into = c("Year", "Month", "Day"), sep = "-")
```

## 3.7. Organize name of columns

```{r}
colnames(injury_data_cleaned)[colnames(injury_data_cleaned) == "Relinquished"] <- "Player_Name"
```


# 4. Data CLean: Player Physical Parameters

## 4.1. Height

```{r}
# Converting Height from Feet-Inches to Centimeters
convert_height_to_cm <- function(height) {
  if (!is.na(height) && grepl("-", height)) {
    parts <- unlist(strsplit(height, "-"))
    feet <- as.numeric(parts[1])
    inches <- as.numeric(parts[2])
    return(round(feet * 30.48 + inches * 2.54, 2))
  } else {
    return(NA)
  }
}

player_data$height_cm <- sapply(player_data$height, convert_height_to_cm)

```


```{r}
# Outputs rows with NA in the height_cm column
na_rows_height <- player_data[is.na(player_data$height_cm), ]
na_rows_height
```

```{r}
# Delete original column
player_data <- player_data[, !colnames(player_data) %in% c("height")]
```


## 4.2. Weight

```{r}
# Convert weight from pounds to kilograms
convert_weight_to_kg <- function(weight) {
  if (!is.na(weight)) {
    return(round(weight * 0.453592, 2))
  } else {
    return(NA)
  }
}
player_data$weight_kg <- sapply(player_data$weight, convert_weight_to_kg)
```


```{r}
# Outputs rows with NA in the weight_kg column
na_rows_weight <- player_data[is.na(player_data$weight_kg), ]
na_rows_weight
```

```{r}
# Delete original column
player_data <- player_data[, !colnames(player_data) %in% c("weight")]
```


## 4.3. Date of birth


```{r}
# Split the date of birth into year, month and day
split_birth_date <- function(birth_date) {
  if (!is.na(birth_date) && grepl(",", birth_date)) {
    parts <- unlist(strsplit(birth_date, " "))
    birth_year <- parts[3]
    birth_month <- parts[1]
    birth_day <- gsub(",", "", parts[2])
    return(data.frame(birth_year = birth_year, birth_month = birth_month, birth_day = birth_day))
  } else {
    return(data.frame(birth_year = NA, birth_month = NA, birth_day = NA))
  }
}
birth_split <- do.call(rbind, lapply(player_data$birth_date, split_birth_date))
player_data <- cbind(player_data, birth_split)
```

```{r}
# Delete original column
player_data <- player_data[, !colnames(player_data) %in% c("birth_date")]
```


## 4.4. Filter NA

```{r}
columns_to_check <- c("year_start", "year_end", "position",
                      "height_cm", "weight_kg",
                      "birth_year", "birth_month", "birth_day")

# Filter rows containing NA
player_data_with_na <- player_data[rowSums(is.na(player_data[columns_to_check])) > 0, ]

# Delete these rows from the original data frame
player_data_cleaned <- player_data[!rowSums(is.na(player_data[columns_to_check])) > 0, ]
```

```{r}
player_data_with_na
```


## 4.5. Check if the position is unique


### 4.5.1. Discovery of repeat name data (players)
```{r}
# Check if each name corresponds to a unique position
name_position_check <- player_data_cleaned %>%
  group_by(name) %>%
  summarise(unique_positions = n_distinct(position)) %>%
  filter(unique_positions > 1)

# View names with multiple positions
print(name_position_check)
```


### 4.5.2. Export repeat name data
```{r}
# Filter out all rows corresponding to these names in the original data
multipositon_players <- player_data_cleaned %>%
  filter(name %in% name_position_check$name)

# View specific rows filtered out
print(multipositon_players)
```

### 4.5.3. Determine whether renames overlap in time

#### 4.5.3.1 Find overlapping data

```{r}
# Define a function to check for overlapping times
check_overlap <- function(years) {
  for (i in 1:(nrow(years) - 1)) {
    if (years$year_end[i] >= years$year_start[i + 1]) {
      return(TRUE)
    }
  }
  return(FALSE)
}
```


```{r}
# Check each name for time overlap
overlap_rows <- multipositon_players %>%
  arrange(name, year_start) %>%                # Sort by name and starting year
  group_by(name) %>%                           # Check each name in groups
  filter(n() > 1) %>%                          # Filtering names with multiple lines
  summarise(overlap = check_overlap(pick(everything()))) %>%  
  # Use pick() instead of cur_data() Use pick() instead of cur_data()
  filter(overlap == TRUE) %>%
  select(name)
```


```{r}
# Output specific rows with time overlap
multipositon_players_time_overlap <- multipositon_players %>%
  filter(name %in% overlap_rows$name)

print(multipositon_players_time_overlap)
```

#### 4.5.3.2. Determine the presence or absence in injury

```{r}
# Check if df1$name exists in df2$name
# df1$name %in% df2$name
multipositon_players_time_overlap$name %in% injury_data_cleaned$Player_Name
```

# 5. Merge Injury and parameter data

## 5.1. Standardize Name Format

```{r}
# Standardize name format: remove extra spaces and standardize to lower case

injury_data_cleaned$Player_Name_Cleaned <- tolower(trimws(injury_data_cleaned$Player_Name))
player_data_cleaned$Player_Name_Cleaned <- tolower(trimws(player_data_cleaned$name))

# Retain only the first two words
injury_data_cleaned$Player_Name_Cleaned <- sapply(
  strsplit(injury_data_cleaned$Player_Name_Cleaned, "\\s+"),
  function(x) paste(x[1:min(2, length(x))], collapse = " ")
)

player_data_cleaned$Player_Name_Cleaned <- sapply(
  strsplit(player_data_cleaned$Player_Name_Cleaned, "\\s+"),
  function(x) paste(x[1:min(2, length(x))], collapse = " ")
)
```


```{r}
# Check for name matches
common_names <- intersect(injury_data_cleaned$Player_Name_Cleaned, player_data_cleaned$Player_Name_Cleaned)
# Outputs the number of names matched
length(common_names)
```


## 5.2. For overlap data

### 5.2.1. Standardization of names

```{r}
multipositon_players_time_overlap$Player_Name_Cleaned <- tolower(trimws(multipositon_players_time_overlap$name))

multipositon_players_time_overlap$Player_Name_Cleaned <- sapply(
  strsplit(multipositon_players_time_overlap$Player_Name_Cleaned, "\\s+"),
  function(x) paste(x[1:min(2, length(x))], collapse = " ")
)
```

### 5.2.2. View name matches

```{r}
# Check for name matches
common_names <- intersect(injury_data_cleaned$Player_Name_Cleaned, multipositon_players_time_overlap$Player_Name_Cleaned)
# Outputs the number of names matched
length(common_names)
```


## 5.3. Remove common_names row from player_data

```{r}
# Retain lines that are not in common_names
player_data_cleaned <- player_data_cleaned[!player_data_cleaned$Player_Name_Cleaned %in% common_names, ]

```


```{r}
# Check for name matches
common_names <- intersect(injury_data_cleaned$Player_Name_Cleaned, player_data_cleaned$Player_Name_Cleaned)
# Output the number of matching names
length(common_names)
```

## 5.4. Merge

### 5.4.1. Convert year data types

```{r}
# Convert injury_data_cleaned$Year to integer type
injury_data_cleaned <- injury_data_cleaned %>%
  mutate(Year = as.integer(Year))
```


### 5.4.2. Merge

```{r}
# Match by name and year range using dplyr's join operation
merged_data <- injury_data_cleaned %>%
  left_join(player_data_cleaned, 
            by = "Player_Name_Cleaned") %>%
  filter(Year >= year_start & Year <= year_end) # %>%
  #select(Year, Month, Day, Team, Player_Name, Notes, Injury_Method, 
         #Injury_Location, Other_Injury, year_start, year_end, position, 
         #height_cm, weight_kg, birth_year, birth_month, birth_day)

```


### 5.4.3. Check the many-to-many

```{r}
# Check the many-to-many
multi_match <- injury_data_cleaned %>%
  left_join(player_data_cleaned, 
            by = "Player_Name_Cleaned") %>%
  filter(!is.na(year_start) & Year >= year_start & Year <= year_end) %>%
  group_by(Player_Name_Cleaned, Year) %>%
  filter(n() > 1)

# Check these question lines
#print(multi_match)

```

## 5.5. Tidy after merging

### 5.5.1. Retain only useful columns

```{r}
merged_data_1 <-  merged_data %>% 
  select(-Player_Name, -name, -college)
```

### 5.5.2. Organize data types

```{r}
merged_data_1$Year <- as.numeric(merged_data_1$Year)
merged_data_1$Month <- as.numeric(merged_data_1$Month)
merged_data_1$Day <- as.numeric(merged_data_1$Day)
```


```{r}
#  merged_data_1$birth_month converted to numeric
merged_data_1$birth_month <- match(merged_data_1$birth_month, month.name)
```

```{r}
merged_data_1$birth_year <- as.numeric(merged_data_1$birth_year)
merged_data_1$birth_month <- as.numeric(merged_data_1$birth_month)
merged_data_1$birth_day <- as.numeric(merged_data_1$birth_day)
```

### 5.5.3. Rename column names

```{r}
merged_data_1 <- merged_data_1 %>%
  rename(
    injury_year = Year,
    injury_month =  Month,
    injury_day = Day 
  )
```

### 5.5.4. Calculate age

```{r}
# Create Date of Birth and Date of Injury columns
merged_data_1$birth_date <- as.Date(with(merged_data_1, paste(birth_year, birth_month, birth_day, sep = "-")), "%Y-%m-%d")
merged_data_1$injury_date <- as.Date(with(merged_data_1, paste(injury_year, injury_month, injury_day, sep = "-")), "%Y-%m-%d")

# Calculate age and round down
merged_data_1$age_at_injury <- floor(as.numeric(difftime(merged_data_1$injury_date, merged_data_1$birth_date, units = "days")) / 365.25)

```

### 5.5.5.. Career (Total｜Before Injury｜After Injury)


```{r}
merged_data_1$career_all <- merged_data_1$year_end- merged_data_1$year_start 
merged_data_1$career_before <- merged_data_1$injury_year- merged_data_1$year_start
merged_data_1$career_after <- merged_data_1$year_end - merged_data_1$injury_year
```

### 5.5.6. Reorder columns

```{r}
merged_data_1 <- merged_data_1 %>%
  select(birth_date,
         injury_date, 
         Player_Name_Cleaned,
         age_at_injury,
         height_cm,
         weight_kg,
         Team,
         position,
         Injury_Method,
         Injury_Location,
         Other_Injury,
         career_before,
         career_after,
         career_all,
         Notes,
         injury_year,
         injury_month,
         injury_day,
         birth_year,
         birth_month,
         birth_day,
         year_start,
         year_end
         )

```




# 6. Player Performance Data

## 6.1. View Basic Information

```{r}
###
###The data frame requires the columns to be of the same data type
###

# Create an empty df to store the results
result <- data.frame(Column = character(0), Data_Type = character(0), stringsAsFactors = FALSE)

# Loop through each column and get the column name and data type
for (col_name in names(All_RegSs_Per_Games)) {
  # Get the data type of the current column
  data_type <- class(All_RegSs_Per_Games[[col_name]])
  
  # Storing results in the results data box
  result <- rbind(result, data.frame(Column = col_name, Data_Type = data_type, stringsAsFactors = FALSE))
}

print(result)
```

```{r}
###Same to previous
###数据框要求列是同一种数据类型
###

# 创建一个空数据框存储结果
result <- data.frame(Column = character(0), Data_Type = character(0), stringsAsFactors = FALSE)

# 循环每一列，获取列名和数据类型
for (col_name in names(All_RegSs_Totals)) {
  # 获取当前列的数据类型
  data_type <- class(All_RegSs_Totals[[col_name]])
  
  # 将结果存储到结果数据框中
  result <- rbind(result, data.frame(Column = col_name, Data_Type = data_type, stringsAsFactors = FALSE))
}

print(result)
```


## 6.2. Basic cleaning

### 6.2.1. Clear Position

```{r}
# Define valid values
valid_positions <- c("C", "PF", "PG", "SF", "SG")

# replace others as NA
All_RegSs_Per_Games$Pos <- ifelse(All_RegSs_Per_Games$Pos %in% valid_positions, 
                                  All_RegSs_Per_Games$Pos, 
                                  NA)

```

### 6.2.2. Drop columns

```{r}
All_RegSs_Per_Games <- All_RegSs_Per_Games %>% 
  select(-Awards)

All_RegSs_Totals <- All_RegSs_Totals %>% 
  select(-Awards)
```


### 6.2.3. De-duplicate lines



```{r}
cat("Before de-duplication, merged_data_1" ,nrow(merged_data_1), "rows.\n")
cat("Before de-duplication, All_RegSs_Per_Games" ,nrow(All_RegSs_Per_Games), "rows. \n")
cat("Before de-duplication, All_RegSs_Totals" ,nrow(All_RegSs_Totals), "rows.\n")
```


```{r}
merged_data_1 <- merged_data_1 %>% distinct()
All_RegSs_Per_Games <- All_RegSs_Per_Games %>% distinct()
All_RegSs_Totals <- All_RegSs_Totals %>% distinct()
cat("After de-duplication, merged_data_1" ,nrow(merged_data_1), "rows.\n")
cat("After de-duplication, All_RegSs_Per_Games" ,nrow(All_RegSs_Per_Games), "rows.\n")
cat("After de-duplication, All_RegSs_Totals" ,nrow(All_RegSs_Totals), "rows.\n")
```


## 6.3. Manipulate name columns

### 6.3.1. Standardize names


```{r}
# Standardize name format: remove extra spaces and standardize to lower case

All_RegSs_Per_Games$Player_Name_Cleaned <- tolower(trimws(All_RegSs_Per_Games$Player))
All_RegSs_Totals$Player_Name_Cleaned <- tolower(trimws(All_RegSs_Totals$Player))

# Retain only the first two words
All_RegSs_Per_Games$Player_Name_Cleaned <- sapply(
  strsplit(All_RegSs_Per_Games$Player_Name_Cleaned, "\\s+"),
  function(x) paste(x[1:min(2, length(x))], collapse = " ")
)

All_RegSs_Totals$Player_Name_Cleaned <- sapply(
  strsplit(All_RegSs_Totals$Player_Name_Cleaned, "\\s+"),
  function(x) paste(x[1:min(2, length(x))], collapse = " ")
)
```

### 6.3.2. Check the number of names

```{r}
common_names_1 <- intersect(merged_data_1$Player_Name_Cleaned, All_RegSs_Per_Games$Player_Name_Cleaned)
common_names_2 <- intersect(merged_data_1$Player_Name_Cleaned, All_RegSs_Totals$Player_Name_Cleaned)
common_names_3 <- intersect(All_RegSs_Per_Games$Player_Name_Cleaned, All_RegSs_Totals$Player_Name_Cleaned)

# Output the number of matching names
cat("merged_data_1 has", length(unique(merged_data_1$Player_Name_Cleaned)), "names.\n")
cat("All_RegSs_Per_Games has", length(unique(All_RegSs_Per_Games$Player_Name_Cleaned)), "names.\n")
cat("All_RegSs_Totals has", length(unique(All_RegSs_Totals$Player_Name_Cleaned)), "names.\n")
cat("merged_data_1 and All_RegSs_Per_Games have", length(common_names_1), "shared names.\n")
cat("merged_data_1 and All_RegSs_Totals have", length(common_names_2), "shared names.\n")
cat("All_RegSs_Per_Games and All_RegSs_Totals have", length(common_names_3), "shared names.\n")
```


## 6.4. Merge of two performance data

### 6.4.1. Columns that are the same

```{r}

# Find identical column names and exclude key columns
key_cols <- c("Player", "Age", "Team")
common_cols <- intersect(setdiff(colnames(All_RegSs_Per_Games), key_cols),
                         setdiff(colnames(All_RegSs_Totals), key_cols))

# Merge the two data frames, and ensure that the key columns are aligned
merged_data <- merge(All_RegSs_Per_Games, All_RegSs_Totals, by = key_cols, suffixes = c("_Games", "_Totals"))

# Compare columns with the same name
comparison_results <- sapply(common_cols, function(col) {
  # Column names with suffixes for both columns
  col_games <- paste0(col, "_Games")
  col_totals <- paste0(col, "_Totals")
  
  # Determine if the columns are the same
  all(merged_data[[col_games]] == merged_data[[col_totals]], na.rm = TRUE)
})


comparison_results

```

### 6.4.2 Select key

#### 6.4.2.1. View Uniqueness
```{r}
# Specify the name of the column to extract
selected_columns <- c("Player", "Age", "Team")  

# Extract as new data frame
new_df <- All_RegSs_Per_Games[, selected_columns, drop = FALSE]

# Check for duplicate rows in the new data frame
has_duplicates <- any(duplicated(new_df))

if (has_duplicates) {
  cat("The new data frame has duplicate rows.\n")
} else {
  cat("The new data frame does not have duplicate rows.\n")
}

# View duplicate rows
duplicate_rows <- new_df[duplicated(new_df), ]
cat("Number of duplicate rows:", nrow(duplicate_rows), "\n")

# Get duplicate rows
duplicate_rows <- new_df[duplicated(new_df) | duplicated(new_df, fromLast = TRUE), ]

# Extract the corresponding rows from the raw data
matching_rows <- merge(All_RegSs_Per_Games, duplicate_rows, by = selected_columns)

print(matching_rows)
```

#### 6.4.2.2. Re-clean rows


```{r}
# Marker league_average Row
matching_rows$is_league_average <- matching_rows$Player == "League Average"

# Group by Player, delete the first row of each group (except League Average)
matching_rows_filtered <- matching_rows %>%
  arrange(Player, Season) %>%  # Group by Player and sort Season in ascending order.
  group_by(Player) %>%
  filter(is_league_average | row_number() > 1) %>%
  ungroup()

# Remove rows from All_RegSs_Per_Games that match matching_rows_filtered
All_RegSs_Per_Games_cleaned <- anti_join(All_RegSs_Per_Games, matching_rows_filtered, 
                                         by = selected_columns)


cat("Number of rows removed:", nrow(All_RegSs_Per_Games) - nrow(All_RegSs_Per_Games_cleaned), "\n")
cat("Number of rows remaining in All_RegSs_Per_Games_cleaned:", nrow(All_RegSs_Per_Games_cleaned), "\n")

```



#### 6.4.2.3. Re-check uniqueness

```{r}
# Extract new data frame
new_df_1 <- All_RegSs_Per_Games_cleaned[, selected_columns, drop = FALSE]

# Check for duplicate rows in the new data frame
has_duplicates_1 <- any(duplicated(new_df_1))

# Print results
if (has_duplicates_1) {
  cat("The new data frame has duplicate rows.\n")
} else {
  cat("The new data frame does not have duplicate rows.\n")
}

# View duplicate rows
duplicate_rows_1 <- new_df_1[duplicated(new_df_1), ]
cat("Number of duplicate rows:", nrow(duplicate_rows_1), "\n")
```

#### 6.4.2.4. Repeat operation on total data


```{r}
# 
new_df_2 <- All_RegSs_Totals[, selected_columns, drop = FALSE]

# 
has_duplicates_2 <- any(duplicated(new_df_2))

# 
if (has_duplicates_2) {
  cat("The new data frame has duplicate rows.\n")
} else {
  cat("The new data frame does not have duplicate rows.\n")
}

# 
duplicate_rows_2 <- new_df_2[duplicated(new_df_2), ]
cat("Number of duplicate rows:", nrow(duplicate_rows_2), "\n")

# 
duplicate_rows_2 <- new_df_2[duplicated(new_df_2) | duplicated(new_df_2, fromLast = TRUE), ]

# 
matching_rows_2 <- merge(All_RegSs_Totals, duplicate_rows_2, by = selected_columns)

# 
print(matching_rows_2)
```

```{r}
# 
matching_rows_2$is_league_average <- matching_rows_2$Player == "League Average"

# 
matching_rows_filtered_2 <- matching_rows_2 %>%
  arrange(Player, Season) %>%  # 
  group_by(Player) %>%
  filter(is_league_average | row_number() > 1) %>%
  ungroup()

# 
All_RegSs_Totals_cleaned <- anti_join(All_RegSs_Totals, matching_rows_filtered_2, 
                                         by = selected_columns)

# 
cat("Number of rows removed:", nrow(All_RegSs_Totals) - nrow(All_RegSs_Totals_cleaned), "\n")
cat("Number of rows remaining in All_RegSs_Totals_cleaned:", nrow(All_RegSs_Totals_cleaned), "\n")

```

```{r}
# 
new_df_3 <- All_RegSs_Totals_cleaned[, selected_columns, drop = FALSE]

# 
has_duplicates_3 <- any(duplicated(new_df_3))

# 
if (has_duplicates_3) {
  cat("The new data frame has duplicate rows.\n")
} else {
  cat("The new data frame does not have duplicate rows.\n")
}

# 
duplicate_rows_3 <- new_df_3[duplicated(new_df_3), ]
cat("Number of duplicate rows:", nrow(duplicate_rows_3), "\n")
```

#### 6.4.2.5. Test for key identity

```{r}
# Sort by name in ascending order and reset row numbers
new_df_1 <- new_df_1 %>%
  arrange(Player) 

# Reset line number
row.names(new_df_1) <- NULL

# Sort by name in ascending order and reset row numbers
new_df_3 <- new_df_3 %>%
  arrange(Player)  # Ascending name order

row.names(new_df_3) <- NULL

#new_df_3 == new_df_1
```

### 6.4.3. Merge!!!

```{r}
# Define the columns to be processed
columns_to_rename <- c("Rk","FG", "FGA", "FG.", "FT", "FTA", "FT.", "TRB", "AST", 
                       "PF", "PTS", "MP", "GS", "ORB", "DRB", 
                       "STL", "BLK", "TOV", "X3P", "X3PA", "X3P.", "X2P", 
                       "X2PA", "X2P.", "eFG.")

# Add suffix to column names of All_RegSs_Totals_cleaned
All_RegSs_Totals_cleaned <- All_RegSs_Totals_cleaned %>%
  rename_with(~ paste0(., "_Ttls"), all_of(columns_to_rename))

# Add suffix to column name of All_RegSs_Per_Games_cleaned
All_RegSs_Per_Games_cleaned <- All_RegSs_Per_Games_cleaned %>%
  rename_with(~ paste0(., "_PrGs"), all_of(columns_to_rename))

```

```{r}
All_RegSs_to_merge <- All_RegSs_Per_Games_cleaned %>% 
  select(-Pos, -G, -Season, -Player_Name_Cleaned)


# Join by key to merge only specified columns
merged_data <- All_RegSs_Totals_cleaned %>%
  left_join(
    All_RegSs_to_merge,
    by = c("Player", "Age", "Team")
  )

print(merged_data)
```

# 7. Merge total data

## 7.1. Unify team names

```{r}
# Creating a mapping table
team_mapping <- data.frame(
  Team_short = c("ATL", "BAL", "BOS", "BRK", "CHA", "CHI", "CLE", "DAL", "DEN", "DET",
                 "GSW", "HOU", "IND", "LAC", "LAL", "MEM", "MIA", "MIL", "MIN", "NOP",
                 "NYK", "OKC", "ORL", "PHI", "PHO", "POR", "SAC", "SAS", "TOR", "UTA",
                 "WAS", "WSB", "SEA", "VAN"),
  Team_full = c("Hawks", "Bullets", "Celtics", "Nets", "Hornets", "Bulls", "Cavaliers", "Mavericks", 
                "Nuggets", "Pistons", "Warriors", "Rockets", "Pacers", "Clippers", "Lakers", 
                "Grizzlies", "Heat", "Bucks", "Timberwolves", "Pelicans", "Knicks", "Thunder", 
                "Magic", "76ers", "Suns", "Blazers", "Kings", "Spurs", "Raptors", "Jazz", 
                "Wizards", "Bullets", "Sonics", "Grizzlies")
)

```


```{r}
# Replace abbreviated team names with full team names
merged_data$Team_full <- team_mapping$Team_full[match(merged_data$Team, team_mapping$Team_short)]

```

## 7.2. merged_data retains only the requirement columns
```{r}
merged_data_2 <- merged_data %>% 
  select(-Pos,
         -Player,
         -Team)
```


## 7.3. Final merging

```{r}
# Perform a merge operation
merged_data_all <- merged_data_1 %>%
  left_join(
    merged_data_2,
    by = c(
      "Player_Name_Cleaned" = "Player_Name_Cleaned",
      "age_at_injury" = "Age",
      "Team" = "Team_full"
    )
  )
```


## 7.4. Write out the data

```{r}
# write.csv(merged_data_all, "merged_data_all.csv", row.names = FALSE)

```









